class: part-slide

```{r, include = FALSE}
opts_chunk$set(eval = FALSE)
```


# Helper<br><i class="fab fa-r-project"></i> Packages

---

# `devtools`: To Develop <i class="fab fa-r-project"></i> Packages

[https://devtools.r-lib.org/](https://devtools.r-lib.org/)

```{r}
# Install devtools from CRAN
install.packages("devtools")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("r-lib/devtools")
```

```{r}
library("devtools")
```

---

# `roxygen2`: To Document the Code

[https://roxygen2.r-lib.org/](https://roxygen2.r-lib.org/)

```{r}
# Install devtools from CRAN
install.packages("roxygen2")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("r-lib/roxygen2")
```

```{r}
library("roxygen2")
```

---

# `roxygen2`: To Document the Code

```{r}
#' Add to message on top of a call to `print()`
#'
#' This function print a greeting message on top of
#' the regular call to `print()`.
#'
#' @inheritParams print
#' @param msg A string to print as a greeting message
#' @export
#' @examples
#' print_msg(1, "The value is: ")
print_msg <- function(x, msg) {
  message(msg)
  x
}
```

---

# `usethis`: To Setup the Environment

[https://usethis.r-lib.org/](https://usethis.r-lib.org/)

```{r}
# Install devtools from CRAN
install.packages("usethis")

# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("r-lib/usethis")
```

```{r}
library("usethis")
```

---

# `usethis`: To Setup the Environment

Global options for `usethis`:

```{r}
options(
  usethis.full_name = "Firstname Lastname",
  usethis.protocol  = "ssh",
  usethis.description = list(
    `Authors@R` = 'person(
      given = "Firstname",
      family = "Lastname",
      role = c("aut", "cre"),
      email = "Firstname.Lastname@cnrs.fr",
      comment = c(ORCID = "0000-0000-0000-0000")
    )',
    Version = "0.0.0.9000"
  )
)
```

---

# `usethis`: To Setup the Environment

Setup Git:

```{r}
use_git_config(
  user.name = "Firstname Lastname", 
  user.email = "Firstname.Lastname@cnrs.fr"
)
# git_sitrep() # To check if Git is configured properly
```

```{bash}
git config --global user.name 'Firstname Lastname'
git config --global user.email 'Firstname.Lastname@cnrs.fr'
git config --global --list
```

---

# `usethis`: To Setup the Environment

First an access token from GitHub needs to be created: https://github.com/settings/tokens

```{r}
browse_github_token() # open the form in a browser
edit_r_environ() # open .Renviron file from your home
```

Edit the `.Renviron` and add your token.
```{txt}
GITHUB_PAT=xloltrglmlm9v0wqb4zgnks86z5uwkqgoat29aar
```

Restart and check if you have access to your token

```{r}
Sys.getenv("GITHUB_PAT")
github_token()
```

---
class: part-slide

# Let's `usethis` <br>Create an <i class="fab fa-r-project"></i> package

---

# `usethis`: A Tidy Package

```{r}
library("usethis")
create_tidy_package()

function (path, name = "RStudio") {
  path <- create_package(path, rstudio = TRUE, open = FALSE)
  old_project <- proj_set(path)
  on.exit(proj_set(old_project), add = TRUE)
  use_roxygen_md()
  use_testthat()
  use_gpl3_license(name)
  use_tidy_description()
  use_readme_rmd(open = FALSE)
  use_lifecycle_badge("experimental")
  use_cran_badge()
  use_cran_comments(open = FALSE)
  use_tidy_github()
  ui_todo("In the new package, remember to do:")
  ui_todo("{ui_code('use_git()')}")
  ui_todo("{ui_code('use_github()')}")
  ui_todo("{ui_code('use_tidy_ci()')}")
  ui_todo("{ui_code('use_pkgdown()')}")
  ui_todo("{ui_code('use_pkgdown_travis()')}")
  proj_activate(path)
}
```

---

# `usethis`: A Tidy Package

Create the package skeleton.

```{r}
path <- create_package("mypackage", rstudio = TRUE, open = FALSE)
```

```{txt}
✔ Creating './'
✔ Creating 'R/'
✔ Writing 'DESCRIPTION'
Package: mypackage
Title: What the Package Does (One Line, Title Case)
Version: 0.0.0.9000
Authors@R (parsed):
    * First Last <first.last@example.com> [aut, cre] (<https://orcid.org/YOUR-ORCID-ID>)
Description: What the package does (one paragraph).
License: What license it uses
Encoding: UTF-8
LazyData: true
✔ Writing 'NAMESPACE'
✔ Writing 'mypackage.Rproj'
✔ Adding '.Rproj.user' to '.gitignore'
✔ Adding '^mypackage\\.Rproj$', '^\\.Rproj\\.user$' to '.Rbuildignore'
```

---

# `usethis`: A Tidy Package

Setup Rstudio project (`rstudio = TRUE` from `create_package()`).

```{r}
old_project <- proj_set(path)
on.exit(proj_set(old_project), add = TRUE)
```

---

# `usethis`: A Tidy Package

Setup roxygen2 as tool to render documentation.

```{r}
use_roxygen_md()
```

```{txt}
✔ Setting Roxygen field in DESCRIPTION to 'list(markdown = TRUE)'
✔ Setting RoxygenNote field in DESCRIPTION to '7.0.2'
● Run `devtools::document()`
```

---

# `usethis`: A Tidy Package

Setup unit test directory with default scripts.

```{r}
use_testthat()
```

```{txt}
✔ Adding 'testthat' to Suggests field in DESCRIPTION
✔ Creating 'tests/testthat/'
✔ Writing 'tests/testthat.R'
● Call `use_test()` to initialize a basic test file and open it for editing.
```

---

# `usethis`: A Tidy Package

Setup licence (several available).

```{r}
use_gpl3_license("mypackage")
```

```{txt}
✔ Setting License field in DESCRIPTION to 'GPL-3 + file LICENSE'
✔ Writing 'LICENSE.md'
✔ Adding '^LICENSE\\.md$' to '.Rbuildignore'
✔ Writing 'LICENSE'
```

---

# `usethis`: A Tidy Package

Format/clean `DESCRIPTION` file.

```{r}
use_tidy_description()
```

Add version for packages listed in `DESCRIPTION` (*e.g.*, `Imports` field).

```{r}
use_tidy_versions()
```

---

# `usethis`: A Tidy Package

Add an Rmarkdown script to render the `README.md` file.

```{r}
use_readme_rmd(open = FALSE)
```

```{txt}
✔ Writing 'README.Rmd'
✔ Adding '^README\\.Rmd$' to '.Rbuildignore'
```

---

# `usethis`: A Tidy Package

Add a development status badge following tidyverse convention.

```{r}
use_lifecycle_badge("experimental")
```

```{txt}
✔ Adding Lifecycle: experimental badge to 'README.Rmd'
● Re-knit 'README.Rmd
```

---

# `usethis`: A Tidy Package

Add a CRAN badge displaying CRAN version.

```{r}
use_cran_badge()
```

```{txt}
✔ Adding CRAN status badge to 'README.Rmd'
● Re-knit 'README.Rmd'
```

---

# `usethis`: A Tidy Package

Add a markdown file (used for the release to CRAN).

```{r}
use_cran_comments(open = FALSE)
```

```{txt}
✔ Writing 'cran-comments.md'
✔ Adding '^cran-comments\\.md$' to '.Rbuildignore'
```

---

# `usethis`: A Tidy Package

Add default files to use in GitHub (*Check them*).  
**Note:** `github/SUPPORT.md` must be edited to fix GitHub repository at line 23.

```{r}
use_tidy_github()
```

```{txt}
✔ Creating '.github/'
✔ Adding '^\\.github$' to '.Rbuildignore'
✔ Writing '.github/CONTRIBUTING.md'
✔ Creating '.github/ISSUE_TEMPLATE/'
✔ Adding '^\\.github/ISSUE_TEMPLATE$' to '.Rbuildignore'
✔ Writing '.github/ISSUE_TEMPLATE/issue_template.md'
✔ Writing '.github/SUPPORT.md'
✔ Writing '.github/CODE_OF_CONDUCT.md'
● Don't forget to describe the code of conduct in your README:
  Please note that the 'mypackage' project is released with a
  [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
  [Copied to clipboard]
```

---

## `usethis`: A Tidy Package

```{r}
ui_todo("In the new package, remember to do:")
ui_todo("{ui_code('use_git()')}")
ui_todo("{ui_code('use_github()')}")
ui_todo("{ui_code('use_tidy_ci()')}")
ui_todo("{ui_code('use_pkgdown()')}")
ui_todo("{ui_code('use_pkgdown_travis()')}")
```

```{txt}
● In the new package, remember to do:
● `use_git()`
● `use_github()`
● `use_tidy_ci()`
● `use_pkgdown()`
● `use_pkgdown_travis()`
```

---

# `usethis`: A Tidy Package

Open the newly created package as an Rstudio project.

```{r}
proj_activate(path)
```
```{txt}
✔ Opening 'C:/mypackage' in new RStudio session
```

---

# `usethis`: A Tidy Package

Setup the package to use Git.

```{r}
use_git()
```

```{txt}
✔ Initialising Git repo
✔ Adding '.Rhistory', '.RData' to '.gitignore'
There are 11 uncommitted files:
* '.github/'
* '.gitignore'
* '.Rbuildignore'
* 'cran-comments.md'
* 'DESCRIPTION'
* 'LICENSE'
* 'LICENSE.md'
* 'mypackage.Rproj'
* 'NAMESPACE'
* 'README.Rmd'
* 'tests/'
Is it ok to commit them?

1: I agree
2: Not now
3: No way
```

---

# `usethis`: A Tidy Package

Setup a repository on GitHub for the package.

```{r}
use_github()
```

```{txt}
✔ Setting active project to 'C:/mypackage'
✔ Checking that current branch is 'master'
● Check title and description
  Name:        mypackage
  Description: What the Package Does (One Line, Title Case)
Are title and description ok?

1: Negative
2: Absolutely not
3: Yup
```

---

# `usethis`: A Tidy Package

Setup continuous integration using travis-CI and code coverage.

```{r}
use_tidy_ci()
```

```{txt}
✔ Writing '.travis.yml'
✔ Adding '^\\.travis\\.yml$' to '.Rbuildignore'
✔ Writing 'codecov.yml'
✔ Adding '^codecov\\.yml$' to '.Rbuildignore'
✔ Adding 'R' to Depends field in DESCRIPTION
✔ Adding 'covr' to Suggests field in DESCRIPTION
✔ Adding '^\\.covrignore$' to '.Rbuildignore'
✔ Adding 'R/deprec-*.R', 'R/compat-*.R' to '.covrignore'
✔ Adding Travis build status badge to 'README.Rmd'
● Re-knit 'README.Rmd'
✔ Adding Codecov test coverage badge to 'README.Rmd'
● Re-knit 'README.Rmd'
✔ Replacing release environment in 'cran-comments.md'
● Turn on travis for your repo at https://travis-ci.org/profile/user
```

---

# `usethis`: A Tidy Package

Add utils for tidy-evaluation (*e.g.*, `.data` pronoun). 

```{r}
use_tidy_eval()
```

```{txt}
✔ Adding 'roxygen2' to Suggests field in DESCRIPTION
✔ Adding 'rlang' to Imports field in DESCRIPTION
✔ Writing 'R/utils-tidy-eval.R'
● Run `devtools::document()`
```

---

# `usethis`: A Tidy Package

Add pipe (`%>%`). 

```{r}
use_pipe()
```

```{txt}
✔ Adding 'magrittr' to Imports field in DESCRIPTION
✔ Writing 'R/utils-pipe.R'
● Run `devtools::document()` to update 'NAMESPACE'
```

---

# `usethis`: A Tidy Package

Add the package `stats` to the package. 

```{r}
use_package("stats")
```

```{txt}
✔ Adding 'stats' to Imports field in DESCRIPTION
● Refer to functions with `stats::fun()`
```

---
class: part-slide

# `devtools`<br>Build / Check / Test

---

# `devtools`: To Build / Check / Test

Frequent development tasks:

* `load_all()` simulates installing and reloading your package, loading <i class="fab fa-r-project"></i> code
  in `R/`, compiled shared objects in `src/` and data files in `data/`.

* `document()` updates generated documentation in `man/`, file collation and `NAMESPACE`.

* `test()` reloads your code with `load_all()`, then runs all `testthat` tests.

* `test_coverage()` runs test coverage on your package with [covr](https://github.com/r-lib/covr). 
  
---

# `devtools`: Build / Check / Test

Building and installing:

* `install()` reinstalls the package, detaches the currently loaded version then reloads the new version with `library()`.

* `build()` builds a package file from package sources. You can use it to build a binary version of your package.


---

# `devtools`: Build / Check / Test

Check and release:

* `check()` updates the documentation, then builds and checks the package locally.  
  `check_win()` checks a package using
  [win-builder](http://win-builder.r-project.org/), and `check_rhub()` checks a package using
  [r-hub](http://log.r-hub.io/). This allows you to easily check
  your package on all systems CRAN uses before submission.

* `release()` makes sure everything is ok with your package (including asking
  you a number of questions), then builds and uploads to CRAN.

---

# Render the `README.md`

Before trying to render the `README.Rmd`, set the chunk with `library(mypackage)` to `eval = FALSE`.

```{r}
rmarkdown::render('README.Rmd',  encoding = 'UTF-8')
```

---

# Update the Code Documentation

```{r}
document()
```

```{txt}
Updating mypackage documentation
Writing NAMESPACE
Loading mypackage
Writing NAMESPACE
Writing pipe.Rd
Writing tidyeval.Rd
```

---

# Check the Package

```{r}
check()
```

```{txt}
-- R CMD check results ---------------------------------------------------------- mypackage 0.0.0.9000 ----
Duration: 19.6s

> checking tests ...
  See below...

-- Test failures ---------------------------------------------------------------------------- testthat ----

> library(testthat)
> library(mypackage)
> 
> test_check("mypackage")
Error: No tests found for mypackage
Execution halted

1 error x | 0 warnings √ | 0 notes √
```
